\name: Combining Bottom-up and top-down approaches through graph learning over interaction networks for drug-target-interaction prediction

### 2021-03-04 paper/main.tex
.

% \subsection{Baseline model}
Furthermore, we hypothesize that some of the predictions made by our
model are the result of the biases in drug--target interaction
datasets. We design a ``na\"ive'' baseline model that predicts the
same list of proteins for each drugs based on the number of
known drug--target interactions for a protein. Formally, given lists
$D$ and $P$ of drugs and proteins and a set of known interactions
$\mathcal{I} := \{(d_i, p_i) \}$, we construct an interaction matrix
$M_{int}\in\{0,1\}^{|D|\times|P|}$ with
\begin{equation*}
	M_{ij} = \begin{cases}
		1 & \text{if } (d_i, p_j)\in \mathcal{I}\\
		0 & otherwise
	\end{cases}
\end{equation*}
describing for all drug--protein pairs whether there is a known
interaction or not. We now rank all proteins $p_j\in P$ descending by
their number of drug interactors by summing over the columns of
$M_{ij}$ and ranking these sums:
\begin{equation*}
	f: P \rightarrow \mathbb{N} \text{ with } f:p_j \mapsto \sum_{i=1}^{|D|}M_{ij}
\end{equation*}
Our ``na\"ive'' predictor $P_k$  predicts all drugs to interact with
the top $k$ targets
% , denoted by $Top_k(P)$ 
with respect to
% $M_{ij}$ from
the introduced ranking:
\begin{equation*}
	P_k: D\times P \rightarrow \{0,1\} \text{ with } P_k: d_i, p_j \mapsto \begin{cases}
		1 & \text{ if }p_j \in Top_{k}(P)\\
		0 & otherwise
	\end{cases}
\end{equation*}
with the only hyperparameter $k$.
The prediction $P_k$ is not dependent on the drug $d_i$ and will
predict the same ranked list of drugs for a given protein $p_j$;
consequently, this na\"ive predictor does not predict any useful
information about interactions between drugs and proteins.
% Also note the possibility for a similar predictor by calculating the top $k$ interacting drugs, respectively.

We apply this naive predictor on both the STITCH and Yamanishi
datasets, using the full dataset as well as 5-fold cross-validation
over drugs and drug--protein pairs.
% Note, that this baseline predictor is not applicable for a protein split CV, as the amount of interactions is unknown. 
For each fold, we gradually increase $k$, eventually saving the best
performance for each fold. Using the full dataset, drug--target split,
and drug split, we obtain \todo[inline]{merge table results in
  text}\todo{are these micro or macro AUC?}.

>>>>

% \subsection{Baseline model}
Furthermore, we hypothesize that some of the predictions made by our
model are the result of the biases in drug--target interaction
datasets. We design a ``na\"ive'' baseline model that predicts the
same list of proteins for each drugs based on the number of
known drug--target interactions for a protein. Formally, given lists
$D$ and $P$ of drugs and proteins and a set of known interactions
$\mathcal{I} := \{(d_i, p_i) \}$, we construct an interaction matrix
$M_{int}\in\{0,1\}^{|D|\times|P|}$ with
\begin{equation*}
	M_{ij} = \begin{cases}
		1 & \text{if } (d_i, p_j)\in \mathcal{I}\\
		0 & otherwise
	\end{cases}
\end{equation*}
describing for all drug--protein pairs whether there is a known
interaction or not. We now rank all proteins $p_j\in P$ descending by
their number of drug interactors by summing over the columns of
$M_{ij}$ and ranking these sums:
\begin{equation*}
	f: P \rightarrow \mathbb{N} \text{ with } f:p_j \mapsto \sum_{i=1}^{|D|}M_{ij}
\end{equation*}
Our ``na\"ive'' predictor $P_k$  predicts all drugs to interact with
the top $k$ targets
% , denoted by $Top_k(P)$ 
with respect to
% $M_{ij}$ from
the introduced ranking:
\begin{equation*}
	P_k: D\times P \rightarrow \{0,1\} \text{ with } P_k: d_i, p_j \mapsto \begin{cases}
		1 & \text{ if }p_j \in Top_{k}(P)\\
		0 & otherwise
	\end{cases}
\end{equation*}
with the only hyperparameter $k$.
The prediction $P_k$ is not dependent on the drug $d_i$ and will
predict the same ranked list of drugs for a given protein $p_j$;
consequently, this na\"ive predictor will not predict any useful
information about interactions between drugs and proteins, but allows
us to estimate the effect of potential biases in our evaluation
dataset.
% Also note the possibility for a similar predictor by calculating the top $k$ interacting drugs, respectively.

We apply this naive predictor on both the STITCH and Yamanishi
datasets, using the full dataset as well as 5-fold cross-validation
over drugs and drug--protein pairs.
% Note, that this baseline predictor is not applicable for a protein split CV, as the amount of interactions is unknown. 
For each fold, we gradually increase $k$ to determine the best
performance for each fold. Using the full dataset, drug--target split,
and drug split, we obtain \todo[inline]{merge table results in
  text}\todo{are these micro or macro AUC? Add to caption!}.
